---
title: Global Warming Potential of Volatile Anaesthetics
output:
    rmarkdown::html_vignette:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{Global Warming Potential of Volatile Anaesthetics}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

# Introduction

The `hypercapnia` package offers functions to estimate the global warming
potential of volatile anaesthetics based on values described in @Ryan2010.

# Data

`hypercapnia` includes the `ianes` data with important key facts about
frequently used anaesthetics.

```{r ianes}
library("hypercapnia")

knitr::kable(ianes)
```

# Reproducing Table 2

In the next code snippet we reproduce the Table 2 of @Ryan2010.
(The values slightly differ because of different precision and ounding errors).

```{r table2}
i <- match(
    rep(c("Sevoflurane", "Isoflurane", "Desflurane"), c(1, 3, 3)),
    ianes$Name
)
fgf <- c(2, rep(c(0.5, 1, 2), 2))
gph <- gram_per_hour(mass = ianes$MW[i], mac = ianes$MAC[i], fgf = fgf)
cde20 <- cde_gram_per_hour(
    mass = ianes$MW[i], mac = ianes$MAC[i], gwp = ianes$GWP20[i], fgf = fgf
)
tbl2 <- data.frame(
    name = ianes$Name[i], mac = ianes$MAC[i] * 100, fgf = fgf,
    gph = gph, gwp20 = ianes$GWP20[i], cde20 = cde20,
    r = cde20 / cde20[1]
)
knitr::kable(
    tbl2,
    col.names = c(
        "Name", "et %", "FGF (L/min)", "Grams/hour", "GWP20", "CDE20 (g/h)",
        "Ratio CDE20"
    ),
    caption = paste0(
        "Comparison of Global Warming Impact of ",
        "Frequently Used Inhaled Anesthetics per MAC-Hour of Use at ",
        "Various Fresh Gas Flows; ",
        "MAC = minimal alveolar concentration; ",
        "et = endtidal;",
        "GWP20 = 20-year global warming potential; ",
        "CDE20 = 20-year carbon dioxide equivalent."
    ),
    digits = 1
)
```

# Reproducing Figure 1

In the next code snippet we reproduce the Figure 1 of @Ryan2010.

```{r figure1}
# order in Ryan2010 Figure 1
o <- c(1, 7, 4)
barplot(
    setNames(tbl2$cde20[o], tbl2$name[o]),
    col = ianes$Color[match(tbl2$name[o], ianes$Name)],
    main = paste0(
        "Relative global warming impact of 1 MAC-hour of ",
        "3 inhaled anesthetics at 2 L fresh gas flow"
    ),
    ylab = expression(CDE[20]),
    sub = "CDE20 = 20-year carbon dioxide equivalent (in grams)"
)
```

# References
